\begin{acknowledgements}
This work would not have been possible to complete without the support and encouragement of many people.

First and foremost, I would like to express my deepest gratitude to my supervisors Prof. Ivan Savenko and Prof. Sergej Flach.
When I first came to this institute, I was not very familiar with the topic of exciton-polariton physics, but they showed great patience in giving me plenty of complicated explanations that helped me to overcome some hard times in this period.
After I started the research, I was inspired by their broad vision about science and their powerful mathematical skills, both analytical and numerical.

I would also like to thank other collaborators in my institute who supported my Ph.D. program.
I had lots of discussions and exciting works with Prof. Daniel Leykam, Prof. Alexei Andreanov, Dr. Kristian Hauser Villegas, Dr. Anton Parafilo, and Dr. Sukjin Yoon.
I learned a lot from them from all our collaborations together.
Thanks to our institute, I had a lot of opportunities to meet and work with other internationl researchers.
I will never forget the edification given by Prof. Yuri Rubo, Prof. Timothy C. H. Liew, Prof. Vadim Kovalev, and Dr. Hugo Flayac, even though we were usually separated by a great distance.
Regarding to this thesis, I must thank Mr. Joel Rasmussen to help me with proofreading.

I started my Ph.D. in the middle of 2016, and during this program, of course, it was not always a pleasant time.
% I must thank my companions, the other Ph.D. students, in my institute.
I must thank my companions, Mr. Ihor Vakulchyk and other Ph.D students in my institute.
We chatted together, played games together, and attended conferences and workshops together.
Without them, I think my memory about my Ph.D. program would be gray and dull.

I also want to thank the Center for Theoretical Physics of Complex Systems at the Institute for Basic Science.
Without their help and financial support, my life in Korea would have been much harder.

At last, I must thank my family for their endless encouragement and unconditional support.
It is their early education in my childhood that lead me to this scientific road.
% This thesis has been done within the Institute for Basic Science, Center for Theoretical Physics of Complex Systems.
\end{acknowledgements}
